Complete Real-Time Liver Model including Glisson Capsule, Vascular Network and Parenchyma

### 2013-02-28 Doc/Miccai2013/miccai2013.tex
paper: intro, intro to experiments
git-svn-id: svn+ssh://scm.gforge.inria.fr/svnroot/sofa-shells@527 7b2761c9-2758-49ef-8c29-9c9f1e1aad1b

\begin{abstract}
This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver (\CD{based on a coupling ?})
We show that the approach is able to reproduce the deformations observed at the local level.
Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

@@

The mechanical importance of the vascular structures in liver is studied in~\cite{peterlik2012}. It shows that that the 
influence of vessels on the mechanical behaviour of the liver is significant, mainly if large deformations occur. 
Since detailed modeling of the vessels would be extremely costly (mainly because of the thickness of the vessel wall 
being less than 250\,{$\mu$}m), the authors propose a composite model allowing for real-time simulation of entire liver
with two vascular trees corresponding to portal and hepatic veins.

The final constituent represented by Glisson's capsule have been already studied in the literature. 
Extensive experiments on porcine liver has been published in~\cite{Sagar}. The measurements show that although being very 
thin (10--20$\mu m$), the Young's modulus of the capsule (reported as $8.22\pm3.42$\,MPa) exceeds the values for the parenchyma by three orders of magnitude,
suggesting the mechanical influence of the membrane on the liver behaviour.

In~\cite{Hollenstein}, a local influence of the capsule has been measured using a special aspiration device. The study was then repeated 
in vivo on human patients during the operation, confirming the mechanical importance of the membrane~\cite{Ahn,Nava}.
To our best knowledge, on attempt has been made to demonstrate the role of the Glisson's capsule on the global behaviour of 
the liver, i.e. during large deformations of the organ.

The main contribution of the paper is twofold: first, we present an efficient real-time model of the liver, taking into account 
the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
linked together via mechanical coupling. We show that the model mimics the local experiments descibed in~\cite{Hollenstein}.

@@

%%Medical simulation is expected to play an integral part in many aspects of
%%medical practice. From diagnosis, through planning or training to computer
%%assisted intervention. A good medical model has to be simple yet efficient
%%to be really usable.
%\CD{develop the applications for which it has to be real-time...}
%
%\CD{First... explain the biomechanics of the liver (for instance: 
%\textit{
%The liver is a complex organ composed by three main structures: the parenchyma (soft tissue, that have a hyper-visco-elastic behavior) is coupled with a dense vascular network (tree-like tubular structure with a stiffer behavior) and is surrounded by the Glisson capsule  (a stiff collagenous membrane).
%It illustrates the need of relying on both volume, surface and line finite elements and the need of proposing coupling methods between them.}
% }
%
%\CD{Yet, many work on liver modeling only consider the parenchyma...}
%The liver is usually modeled as a homogeneous isotropic (visco-)elastic
%material \TG{refs}.% It is however non-homogeneous and modeling it as
%%homogeneous necessarily requires complex non-linear material descriptions.
%Recently
%Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
%structures on the liver deformation.
%
%Another often overlooked component is the Glisson's capsule.
%Biomechanical measurements report that the capsule has very small
%thickness (approx. 20 $\mu$m for porcine liver \cite{Umale2011}).
%On the other hand it has in three orders of magnitude higher stiffness than
%the parenchyma.
%Because of this it plays an important role in local
%deformations of the liver \cite{Ahn2010,Hollenstein2006}.
%It was reported that evaluation of the liver as a homogeneous material
%without capsule can lead to overestimation of local deformations up to the factor of 3
%\cite{Hollenstein2006}.
%
%While the influence on local deformation was reported, no experimental
%results show whether the capsule also plays an important role on the global
%deformations of the liver. We propose a composite model that models the
%capsule with membrane FEM elements. We show that the model is able to
%simulate expected local behaviour reported in literature. Further more we
%show that the capsule plays also an important role in the deformation on
%global scale.
%
\TG{summary of following sections.}

@@

To evaluate the model we have implemented in the
SOFA\footnote{www.sofa-framework.org} framework and performed a set of
numerical simulations. First local deformations were compared to the
results reported in literature to validate the method. Then the model of
the complete liver was subjected to natural global deformations to asses
the importance of the capsule on the deformation.

Material properties used in the simulations were for a porcine liver as
reported by Umale \cite{Umale2013}. The
parenchyma was modeled with elastic modulus 1980~Pa and the capsule has
elastic modulus 8~MPa and thickness 20~$\mu$m.
\TG{values for blood vessels}

\CD{If the experiments demonstrate that the stiffness of the material is different, the thickness is very very small... so we have to evaluate if it is really necessary to include the model of the capsule for both local and global deformations.}
\TG{I didn't undersand that. What do you mean?}

@@

\end{document}
% vim:set et sw=2 tw=75 fdm=marker fdl=3 fdc=4 isk+=_,-:

>>>>

\begin{abstract}
This work provides a complete physical model for the deformations of the liver with its three constituents: parenchyma, vessels and the Glisson's capsule.
Some recent experimental results \cite{Ahn2010} have shown that there is a significant difference of stiffness between the Glisson's capsule and the parenchyma. 
Moreover, it has been observed that the capsule plays an important role for the deformation at the local level \cite{Hollenstein2006}.
However, the Glisson's capsule is very thin and its influence for the global deformations of the liver was never highlighted.
In this work, we propose to use a membrane FEM model for the capsule, that is coupled with a vascularized model of the liver.
We show that the approach is able to reproduce the deformations observed at the local level.
Additionally, we show, by simulating natural global deformations of the liver observed on real images (due to gravity), that the results have important differences with and without the capsule.

@@

The mechanical importance of the vascular structures in liver is studied in~\cite{Peterlik2012}. It shows that that the 
influence of vessels on the mechanical behaviour of the liver is significant, mainly if large deformations occur. 
Since detailed modeling of the vessels would be extremely costly (mainly because of the thickness of the vessel wall 
being less than 250\,{$\mu$}m), the authors propose a composite model allowing for real-time simulation of entire liver
with two vascular trees corresponding to portal and hepatic veins.

The final constituent represented by Glisson's capsule have been already studied in the literature. 
Extensive experiments on porcine liver has been published in~\cite{Umale2013}. The measurements show that although being very 
thin (10--20$\mu m$), the Young's modulus of the capsule (reported as $8.22\pm3.42$\,MPa) exceeds the values for the parenchyma by three orders of magnitude,
suggesting the mechanical influence of the membrane on the liver behaviour.

In~\cite{Hollenstein2006}, a local influence of the capsule has been measured using a special aspiration device. The study was then repeated 
in vivo on human patients during the operation, confirming the mechanical importance of the membrane~\cite{Ahn2010,Nava2008}.
To our best knowledge, on attempt has been made to demonstrate the role of the Glisson's capsule on the global behaviour of 
the liver, i.e. during large deformations of the organ.

The main contribution of the paper is twofold: first, we present an efficient real-time model of the liver, taking into account 
the parenchyma, vascularization and Glisson's capsule. The model is based on three different finite element representations for each constituent,
linked together via mechanical coupling. We show that the model mimics the local experiments descibed in~\cite{Hollenstein2006}.

@@

%Medical simulation is expected to play an integral part in many aspects of
%medical practice. From diagnosis, through planning or training to computer
%assisted intervention.
%Medical training application, like the ones for training laparoscopic
%surgery, have to operate in real-time and a good medical model has to be
%simple yet efficient to be really usable in this environment.
%
%Liver is largest internal organ formed mostly by a soft tissue (parenchyma)
%that is hyper-elastic and also visco-elastic. The parenchyma is interwoven
%by dense vascular network, this tree-like tubular structure is stiffer then
%the parenchyma. Finally the parenchyma is covered by very stiff collagenous
%membrane that preserves the integrity of the organ.
%
%Strangely, most of the work on liver modelling only considers the
%parenchyma and model the liver as single homogeneous object \TG{Refs}.
%Only recently
%Peterl\'{i}k et al. \cite{Peterlik2012} showed the importance of the vascular
%structures on the liver deformation.
%
%Another overlooked component is the Glisson's capsule.
%Biomechanical measurements report that the capsule has very small
%thickness (approx. 20 $\mu$m for porcine liver \cite{Umale2011}).
%On the other hand it has in three orders of magnitude higher stiffness than
%the parenchyma.
%Because of this it plays an important role in local
%deformations of the liver \cite{Ahn2010,Hollenstein2006}.
%It was reported that evaluation of the liver as a homogeneous material
%without capsule can lead to overestimation of local deformations up to the factor of 3
%\cite{Hollenstein2006}.
%
%While the influence on local deformation was reported, no experimental
%results show whether the capsule also plays an important role in global
%deformations of the liver. We propose a composite model that models the
%capsule with membrane FEM elements. We show that the model is able to
%simulate expected local behaviour reported in literature. Further more we
%show that the capsule plays also an important role in the deformation on
%global scale.

\TG{summary of following sections.}

@@

To evaluate the model it was implemented in the
SOFA\footnote{www.sofa-framework.org} framework and a set of
numerical simulations was performed. First local deformations were compared to the
results reported in literature to validate the method.

To show that, despite its very small thickness, the membrane cannot be
neglected even in the context of global deformations and its overall
stiffness plays an important role, the model of
the complete liver was subjected to natural global deformations.

In the simulations we used material properties for a porcine liver as
reported by Umale \cite{Umale2013}. The
parenchyma was modeled with elastic modulus 1980~Pa and the capsule has
elastic modulus 8~MPa and thickness 20~$\mu$m.
\TG{values for blood vessels}

@@

\end{document}
% spell: setl spell spelllang=en spellfile=spelldict.en.utf-8.add
% vim:set et sw=2 tw=75 fdm=marker fdl=3 fdc=4 isk+=_,-:
